Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-14 journal paper/content/sec-theory.tex
Revised HMR subsection.

\newcommand{\BIRgr}{\BIR^{\texttt{gr}}}

Next, let us argue that \alg[2] is monotone. We spell out the proof for the sake of intuition. For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2], and let $\BIRgr_1(n)$ be the $\BIR$ of the ``greedy choice" after $n-1$ steps of \alg[1]. Let 
    $\mathtt{reg}_2(n) = n\cdot \max_a \mu_a - \rew_2(n)$ 
be the (frequentist) instantaneous regret of \alg[2]. Then by definition of the greedy modification it holds that 
\begin{align}\label{eq:random-BIR}
\E\sbr{ \mathtt{reg}_2(n) \mid M_n = m}
    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m) .\\
\BIR_2(n)
    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr_1(M_n) }.
\end{align}
Now, monotonicity of $\BIR_2$ follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$. The former is assumed, and the latter is the monotonicity-in-information property of the ``greedy step" (see \ascomment{pointer} for a precise formulation and proof).

@@

\begin{corollary}\label{cor:random-greedy}
Posit a symmetric \HardMaxRandom response function. Suppose both principals can only choose from monotone MAB algorithms. Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy. Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
\asedit{These conclusions hold even if the principals' strategy sets are any collections of monotone MAB algorithms that are closed under greedy mixing.}%
\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains all greedy mixtures of all algorithms in this set.}
\end{corollary}

\ascomment{Steven: I revised this example to clarify what it is supposed to mean. But the conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$! Perhaps what we derive is that
$\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }

For the sake of intuition, let us show how \alg[1] can be a better algorithm than its greedy modification, in the sense of having a strictly smaller \BIR.

\begin{example}
Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}.
Let $\BIRgr_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
\begin{align}\label{eq:cor:random-greedy-ex-assumption}
 \tfrac{1}{c}\; \BIRgr_1(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
 \qquad \text{for all local steps $n$}.
\end{align}
Then, for a sufficiently small $p>0$, we have $\BIR_2(n)>\BIR_1(n)$ for all local steps $n\geq n_0$.
\end{example}

\begin{proof}
By definition of the modified algorithm we have that
\begin{align}\label{eq:random-BIR}
\BIR_2(n)
    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m)}.
\end{align}
In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.

>>>>

\newcommand{\gr}{\texttt{gr}}
\newcommand{\BIRgr}{\BIR^{\gr}}
\newcommand{\alggr}{\alg[\gr]}

Let us analyze the \BIR of the greedy modification (we will use it for what follows). Use \alg[1] and \alg[2] as in the theorem. For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2]. Let \alggr be a hypothetical algorithm which at each step $n$ outputs the ``Bayesian-greedy choice" based on the data collected by \alg[1] in the first $n-1$ steps. Let $\BIRgr(n)$ be the \BIR of this algorithm. Let
    $\mathtt{reg}_2(n) = n\cdot \max_a \mu_a - \rew_2(n)$
be the (frequentist) instantaneous regret of \alg[2]. Then 
\begin{align}
\E\sbr{ \mathtt{reg}_2(n) \mid M_n = m}
    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m). \nonumber \\
\BIR_2(n)
    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr(M_n) }.
    \label{eq:BIR-modification}
\end{align}

\ascomment{Steven: do you buy this example? }

Let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2]. We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data. Formally, we posit a convex function $f(\cdot)$ and $m_0>0$ such that
\[ \BIRgr(n) \geq f(n) 
    \quad\text{and}\quad 
    f\rbr{\E[M_n]} \geq 2\, \BIR_1(n) \quad \forall n\geq m_0 .  \]
Then by \eqref{eq:BIR-modification} and Jensen's inequality, for each step $n\geq m_0$ we have
\[\BIR_2(n) 
    \geq  \E\sbr{ \BIRgr(M_n) }
    \geq \E\sbr{ f(M_n) } 
    \geq f\rbr{ \E[M_n] }
    \geq 2\,\BIR_1(n). \]

Second, let us argue that \alg[2] is monotone. This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are monotone. The latter follows from a ``monotonicity-in-information" property of the ``greedy step": essentially, it can only get better with more information
(see \ascomment{add a pointer}).

@@

\begin{corollary}\label{cor:random-greedy}
Posit a symmetric \HardMaxRandom response function. Suppose both principals can only choose from monotone MAB algorithms. Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy. Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
\asedit{These conclusions hold even if each principal's strategy set is a subset of monotone MAB algorithms that is closed under the greedy modification.}%
\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
\end{corollary}

\ascomment{Steven: OLD EXAMPLE. I revised it to clarify what it is supposed to mean. But my new example is better, isn't it?}

\ascomment{The conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$! Perhaps what we derive is that
$\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }

\begin{example}
Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}.
Let $\BIRgr(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
\begin{align}\label{eq:cor:random-greedy-ex-assumption}
 \tfrac{1}{c}\; \BIRgr(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
 \qquad \text{for all local steps $n$}.
\end{align}
Then, for a sufficiently small $p>0$, we have $\BIR_2(n)>\BIR_1(n)$ for all local steps $n\geq n_0$.
\end{example}

\begin{proof}
By definition of the modified algorithm we have that
\begin{align}\label{eq:random-BIR}
\BIR_2(n)
    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m)}.
\end{align}
In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.
